Model-based pretreatment for regression with correlated variables.

### 2014-04-01 www/article13.tex
up structure
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@493 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

	\section{Estimating structure of sub-regressions with a Markov chain}	
\subsection{Detailed generative model}
	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 
	\begin{equation}
			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
		\end{equation}
		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed. Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.

		Variables in $X_1$ are supposed to be independent.
	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways. %compute corresponding $\psi$ according to it. %and then improve the walk (will keep a structure only if it is really relevant).%and give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
	We now have a full generative model.
	\subsection{Structure comparison} \label{compstruct}
  To obtain the marginal regression model we need $S$, the linear structure between the covariates.
	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}. We will prefer this kind of comparison criterion instead of cross-validation that is very time-consuming and thus not friendly with combinatory problematics.
	We note $\Theta$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|X)&\propto& P(X|S)P(S) \\
		-2\log P(X|S)&\approx & BIC=-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n)\\  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
	But we can imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context. 
\\	
		We will now use the following notation: $\psi(S)=\psi(X|S)$	where $\psi(X|S)$ is the chosen criterion, that will be $BIC$ with hierarchical uniform hypothesis in our numerical results. We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.

	\subsection{The neighbourhood}
	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
	\\
	For each step, starting from $S \in \mathcal{S}$ we define a neighbourhood:
		\begin{eqnarray}
		\mathcal{V}_{S,j}&=& \{S \}\cup \{ S^{(i,j)} |1\leq i \leq p, i\neq j  \} \\
		\textrm{where }\ \ j &\sim & \mathcal{U}(\{1,\dots,p\}) 
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item else (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}

@@

	\subsection{The walk}
	\subsubsection{Transition probabilities}
	We first make the approximation
	\begin{equation}
		P(S|X)\approx exp(\psi(S)).
	\end{equation}
	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
	At each step the markov chain moves with probabiliy:
	\begin{eqnarray}
			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2} \psi(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}\psi(S_l))} \\
	\end{eqnarray}
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité

>>>>

\section{Sub-regressions model selection}	
	\subsection{Modeling the uncorrelated covariates: a full generative approach}
	To be able to compare structures with probabilistic criterions, we need a full generative model. We suppose that variables in $X_1$ follow Gaussian mixtures. 
	\begin{equation}
			\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)=\mathcal{GM}(\pi_j;\mu_j;\sigma^2_j) \textrm{ with } \pi_j,\mu_j,\sigma^2_j \textrm{ vectors of size } k_j; \label{mixtureX1}
		\end{equation}
		The great flexibility \cite{mclachlan2004finite} of such models makes our model more robust but one can use other laws if needed. Gaussian case is just a special case ($k_j=1$) of Gaussian mixture so it is included in our hypothesis.

		Variables in $X_1$ are supposed to be independent.
	Thus if one have some hypothesis on the distribution of some variables (exponentially distributed for example) it is possible to use it without impacting the model in other ways. %compute corresponding $\psi$ according to it. %and then improve the walk (will keep a structure only if it is really relevant).%and give it as an input of \textsc{CorReg} and then improve the efficiency of the algorithm (it will find a structure only if it is really relevant).
	We now have a full generative model.
	\subsection{Penalization of the integrated likelihood} \label{compstruct}
  To obtain the marginal regression model we need $S$, the linear structure between the covariates.
	Our full generative generative model allows us to compare structures with criterions like the Bayesian Information Criterion ($BIC$) which penalize the log-likelihood according to the complexity of the structure~\cite{BIChuard}. We will prefer this kind of comparison criterion instead of cross-validation that is very time-consuming and thus not friendly with combinatory problematics.
	We note $\Theta$ the set of the parameters of the generative model
	\begin{eqnarray}
		P(S|X)&\propto& P(X|S)P(S) \\
		-2\log P(X|S)&\approx & BIC=-2\mathcal{L}(X,S,\Theta)+|\Theta|\log(n)\\  
	\end{eqnarray}
	But $BIC$ tends to give too complex structures because we test a great range of models. 
	Thus we choose to penalise the complexity a bit more with a hierarchical uniform {\it a priori} law $P(S)=P(I_1 | p_1,I_2,p_2)P(p_1|I_2,p_2)P(I_2|p_2)P(p_2)$  instead of a simple uniform law on $S$.
%	 Thus we have :
%		\begin{eqnarray}
%		BIC^*(X|S)&=&BIC(X|S) +\ln(P(S)) \label{Bicstar}
%	\end{eqnarray}		
	It increases penalty on complexity for $p_2<\frac{p}{2}$ and $p_1^j<\frac{p}{2}$ . Hence %when using $BIC*$ 
	this constraint on $\hat{p}_2$ and $\hat{p}_1^j$ is made in the MCMC in the followings.
	But we can imagine to use other criterions, like the $RIC$ (Risk Inflation Criterion \cite{foster1994risk}) that choose a penalty in $\log p$ instead of $\log n$ and thus gives more parsimonious models when $p$ is larger than $n$ (high dimension) or any other criterion \cite{george1993variable} thought to be better in a given context. 
\\	
		We will now use the following notation: $\psi(S)=\psi(X|S)$	where $\psi(X|S)$ is the chosen criterion, that will be $BIC$ with hierarchical uniform hypothesis in our numerical results. We do not change $BIC$ but only $P(S)$ so the properties are the same as classical $BIC$ but we will obtain better results when our hypothesis is verified.

\subsection{MCMC algorithm}
	\subsubsection{The neighbourhood}
	Let's define $\mathcal{S}$ the ensemble of feasible structures (those with $I_1\cap I_2=\emptyset$).
	\\
	For each step, starting from $S \in \mathcal{S}$ we define a neighbourhood:
		\begin{eqnarray}
		\mathcal{V}_{S,j}&=& \{S \}\cup \{ S^{(i,j)} |1\leq i \leq p, i\neq j  \} \\
		\textrm{where }\ \ j &\sim & \mathcal{U}(\{1,\dots,p\}) 
	\end{eqnarray}	
	With $S^{(i,j)}$ defined by the following algorithm :
	\begin{itemize}
		\item if $i \notin I_i^j$ (add): 
			\begin{itemize}
				\item $I_1^j=I_1^j\cup \{i\}$
				\item $I_1^i=\emptyset$ (explicative variables can't depend on others : column-wise relaxation)
				\item $I_1=I_1 \setminus \{j\}$ (dependent variables can't explain others : row-wise relaxation) 
			\end{itemize}			 
		\item else (remove): $I_1^j=I_1^j\setminus \{i\}$
	\end{itemize}

@@

	\subsubsection{Transition probabilities}
	We first make the approximation
	\begin{equation}
		P(S|X)\approx exp(\psi(S)).
	\end{equation}
	The algorithm follows a time-homogeneous markov chain whose transition matrix $\mathcal{P}$ has $|\mathcal{S}|$ rows and columns (combinatory so we'll just compute the probabilities when we need them).
	At each step the markov chain moves with probabiliy:
	\begin{eqnarray}
			\forall (S,\tilde{S}) \in \mathcal{S}^2 : \mathcal{P}(S,\tilde{S})&=&\sum_{j=1}^p \mathbf{1}_{ \{\tilde{S}\in \mathcal{V}_{S,j}\} }\frac{\exp(-\frac{1}{2} \psi(\tilde{S}))}{\sum_{S_l\in \mathcal{V}_{S,j}}\exp(-\frac{1}{2}\psi(S_l))} \\
	\end{eqnarray}
	And $\mathcal{S}$ is a finite state space.%la relaxation rend P non symétrique mais ne remets  pas en cause l'homogénéité
